### 2022-10-05 targeted/output_control.Rmd
Now tracking targeted data

---
title: "Peak area control doc"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options("readr.num_columns" = 0)

library(tidyverse)

already_good <- 0.1
min_improvement <- 0.4

cv <- function(x)sd(x)/mean(x)

clean_stans <- read.csv("../metadata/made_data/clean_stans.csv")
name_conversion <- clean_stans %>%
  select(new=compound_name, old=compound_name_old) %>%
  distinct()
```

# Convert raw data ----

First step is getting the data out of Skyline/MSDIAL. This is doable by script
if the original .sky and .mdf files are available, as they should be in this
case in the raw_data folder broken down by cruise.

We also need the console interfaces to both Skyline and MSDIAL.
MSDIAL's comes with the app and is pointed to with a raw file path to wherever
the MSDIAL files were extracted, and Skyline requires a separate download of
their command line version.

Mesotransect was integrated with MSDIAL, Falkor and Mesocenter used Skyline.

```{r grabRAnotes}
if(!file.exists("raw_data/Raafay_HILIC_notes.csv")){
  library(googlesheets4)
  if(!gs4_has_token())stop("Refresh Google token")
  # googlesheets4::gs4_auth()

  manual_pos <- read_sheet(ss = "1lU7g5rfem81s_0fqMP93yzegogCN29JJEdsIESfAX7c", 
                           skip = 1, sheet = "HILIC_POS") %>%
    select(compound_name=`Compound name`,
           integrated=`Integrated? (Y/N)`,
           in_samps=`In samples? (Y/N)`) %>%
    mutate(polarity="pos")
  manual_neg <- read_sheet(ss = "1lU7g5rfem81s_0fqMP93yzegogCN29JJEdsIESfAX7c",
                           sheet = "HILIC_NEG") %>%
    select(compound_name=`Compound name`,
           integrated=`Integrated? (Y/N)`,
           in_samps=`In samples? (Y/N)`) %>%
    mutate(polarity="neg")
  manual_quality <- rbind(manual_pos, manual_neg) %>%
    mutate(compound_name=str_replace(compound_name, "_ ", ", ")) %>%
    filter(!is.na(compound_name)) %>%
    filter(compound_name!="Internal Standards") %>%
  left_join(name_conversion, by=c(compound_name="old")) %>%
  mutate(compound_name=new) %>%
  select(-new)
  write.csv(manual_quality, file = "raw_data/Raafay_HILIC_notes.csv", 
            row.names = FALSE)
}
manual_quality <- read.csv("raw_data/Raafay_HILIC_notes.csv") 
```

SkylineRunner.exe output relating to exporting files:
Exporting reports
ÚÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÂÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄ¿
³--report-name=<name>          ³The name of a report to export as it appears ³
³                              ³in the Skyline Export Report form.           ³
ÃÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÅÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄ´
³--report-file=path/to/file.csv³The path to export the report to. Required if³
³                              ³--report-name is specified.                  ³
ÃÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÅÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄ´
³--report-add=path/to/file.skyr³Add reports from a saved report template file³
³                              ³to the running program.                      ³
ÃÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÅÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄ´
³--report-conflict-resolution= ³Specifies how to resolve report name         ³
³<overwrite | skip>            ³conflicts, by either overwriting or skipping ³
³                              ³them, when using --report-add (default is to ³
³                              ³output an error message for conflicts)       ³
ÃÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÅÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄ´
³--report-format=<csv | tsv>   ³CSV for comma-separated reports (or semicolon³
³                              ³separated, depending on your localization) or³
³                              ³TSV for tab separated reports [default CSV]  ³

```{r Falkor Skyline data extraction}
FK_POS_cmd <- paste(
  "SkylineRunner.exe",
  "--in=raw_data/Falkor/HILIC_QE_POSITIVE_Falkor_1.sky", 
  '--report-name="Ingalls_Lab_QE_Transition Results"',
  '--report-file="raw_data/Falkor/Ingalls_Lab_QE_Transition Results_POS.csv"',
  '--report-format="csv"'
)
system(FK_POS_cmd)

FK_neg_cmd <- paste(
  "SkylineRunner.exe",
  "--in=raw_data/Falkor/HILIC_QE_NEGATIVE_Falkor.sky", 
  '--report-name="Ingalls_Lab_QE_Transition Results"',
  '--report-file="raw_data/Falkor/Ingalls_Lab_QE_Transition Results_NEG.csv"',
  '--report-format="csv"'
)
system(FK_neg_cmd)
```

```{r Mesocenter Skyline data extraction}
MC_POS_cmd <- paste(
  "SkylineRunner.exe",
  "--in=raw_data/Mesocenters/HILIC_QE_POSITIVE_Wei_HRM_DepthProfile.sky", 
  '--report-name="Ingalls_Lab_QE_Transition Results"',
  '--report-file="raw_data/Mesocenters/Ingalls_Lab_QE_Transition Results_POS.csv"',
  '--report-format="csv"'
)
system(MC_POS_cmd)

FK_NEG_cmd <- paste(
  "SkylineRunner.exe",
  "--in=raw_data/Mesocenters/HILIC_QE_NEGATIVE_Wei_HRM_DepthProfile.sky", 
  '--report-name="Ingalls_Lab_QE_Transition Results"',
  '--report-file="raw_data/Mesocenters/Ingalls_Lab_QE_Transition Results_NEG.csv"',
  '--report-format="csv"'
)
system(FK_NEG_cmd)
```

```{r MSDIAL autoexport things not yet}
# Not implemented yet, check on thread below for updates
# http://www.metabolomics-forum.com/index.php?topic=1820.0
# Instead, open an existing .mtd2 file, click the "Export" tab, select "Alignment result", specify the output directory as "Area_HILICPos_EddyTransect.csv", check a couple of the boxes (Raw Data Area is the only critical one), change the output format to msp(?), and hit the "Export" button at the bottom. You then need to open this file in Excel and save as CSV.

# Haven't been able to find the raw data for negative mode so I'm currently just copying this off the Google Drive (G:\Shared drives\Ingalls Lab\Collaborative Projects\MESO-SCOPE\EddyTransect\HILIC_Neg_MSDial)
```

# Falkor things

Start with the Falkor data, since it's nice and neat. Here we're pulling from
the raw_data/Falkor folder, which has the Skyline files, their output, and
will soon have Raafay's notes on peak quality.

These CSV files were obtained from Skyline by exporting the transition results,
with a "_POS" or "_NEG" added to distinguish them.

```{r falkorSkylineImport}
# Read in the data and clean it up a little bit
manual_pos <- read.csv("raw_data/Falkor/Ingalls_Lab_QE_Transition Results_POS.csv") %>%
  mutate(polarity="pos")
manual_neg <- read.csv("raw_data/Falkor/Ingalls_Lab_QE_Transition Results_NEG.csv") %>%
  mutate(polarity="neg")
manual_ints <- rbind(manual_pos, manual_neg) %>%
  select(filename=Replicate.Name, compound_name=Precursor.Ion.Name, 
         area=Area, polarity) %>% 
  mutate(filename = ifelse(str_detect(filename, "DDA"),
                           str_replace(filename, "Poo", "DDA"),
                           filename)) %>%
  mutate(filename = ifelse(str_detect(filename, "QC"),
                           str_replace(filename, "Poo", "QC"),
                           filename)) %>% 
  mutate(compound_name=str_replace(
    compound_name, "Monophosphate, 15N5", "monophosphate, 15N5")) %>%
  left_join(name_conversion, by=c("compound_name"="old")) %>% 
  mutate(compound_name=new) %>% 
  select(-new) %>% 
  left_join(manual_quality, by=c("compound_name", "polarity"))

# Filter for only those peaks that Raafay actually integrated
clean_ints_falkor <- manual_ints %>%
  filter(integrated=="Y"&in_samps%in%c("Y", "Maybe?", "Kinda?", "Barely", "Kinda")) %>%
  select(-integrated, -in_samps) %>% 
  mutate(area=as.numeric(area))
```

```{r falkorBMIS}
pooled <- clean_ints_falkor %>%
  filter(str_detect(filename, "_Poo_"))

pooled_IS <- pooled %>%
  left_join(clean_stans %>% select(compound_name, compound_type) %>% distinct(), by=c("compound_name")) %>% 
  filter(compound_type=="Internal Standard") %>%
  select(-compound_type, -polarity) %>%
  rename_with(~paste0("IS_", .), -filename) %>%
  add_row(filename=unique(.$filename), IS_compound_name="None", IS_area=1)

initial_CVs <- pooled %>%
  mutate(area=ifelse(str_detect(filename, "Half"), area*2, area)) %>%
  group_by(compound_name) %>%
  summarize(init_CV=cv(area))

standardized_CVs <- full_join(pooled, pooled_IS, by=c("filename")) %>%
  group_by(compound_name, IS_compound_name) %>%
  summarize(norm_CV=cv(area/IS_area), .groups="drop")

chosen_BMIS_falkor <- standardized_CVs %>%
  left_join(initial_CVs) %>%
  group_by(compound_name) %>%
  mutate(MIS_improvement=(init_CV - norm_CV)/init_CV) %>% 
  filter(MIS_improvement>min_improvement|IS_compound_name=="None") %>%
  arrange(desc(MIS_improvement)) %>%
  group_by(compound_name) %>%
  slice(1) %>%
  select(compound_name, BMIS=IS_compound_name)

BMISed_falkor <- clean_ints_falkor %>%
  left_join(chosen_BMIS_falkor, by="compound_name") %>%
  left_join(clean_ints_falkor %>% rename(IS_area=area) %>% select(-polarity), 
            by=c(BMIS="compound_name", "filename")) %>%
  group_by(compound_name) %>%
  mutate(BMISed_area=area/IS_area*mean(IS_area)) %>%
  mutate(BMISed_area = ifelse(BMIS=="None", area, BMISed_area)) %>%
  mutate(BMISed_area = ifelse(str_detect(filename, "Half")&BMIS=="None", 
                              BMISed_area*2, BMISed_area)) %>%
  select(-IS_area)
```

```{r falkorRFs}
falkor_rundate <- strptime("190715", "%y%m%d")
falkor_stans <- clean_stans %>%
  filter(date_added<=falkor_rundate)

rfs_falkor <- falkor_stans %>%
  left_join(BMISed_falkor, by=c("compound_name", "polarity")) %>%
  filter(str_detect(filename, "_Std_")) %>%
  filter(compound_type!="Internal Standard") %>%
  select(-compound_type) %>%
  mutate(mix=paste0("H2OinMatrix|", paste0(mix, "InMatrix"), "|", paste0(mix, "InH2O"))) %>%
  filter(str_detect(filename, mix)) %>%
  mutate(type=str_replace(str_extract(filename, mix), "Mix[12]", "Mix")) %>%
  select(-mix) %>%
  group_by(compound_name, type) %>%
  summarize(mean_area=mean(area, na.rm = TRUE), conc_um=unique(conc_um)) %>%
  pivot_wider(names_from = type, values_from=mean_area) %>%
  mutate(rf_h2o=MixInH2O/conc_um) %>%
  mutate(rf_matrix=(MixInMatrix - H2OinMatrix)/conc_um) %>%
  mutate(rf_ratio=rf_matrix/rf_h2o)

# 400uL vial, 10L sample, diluted x2
falkor_data <- rfs_falkor %>% 
  left_join(BMISed_falkor, by="compound_name") %>%
  left_join(clean_stans) %>%
  mutate(uM = BMISed_area/rf_h2o * 0.0004/10 * 1/rf_ratio * 2) %>% 
  mutate(nM=uM*1000) %>%
  select(compound_name, nM, filename, polarity)
```

# Mesotransect things

Continue with the MESOSCOPE transect data. Here again, we use Raafay's notes
on peak quality to remove the gross peaks.

```{r transectSkylineImport}
manual_pos <- read_csv("raw_data/Mesotransect/Area_HILICPos_EddyTransect.csv", skip = 4) %>%
  filter(`Metabolite name`!="Unknown") %>%
  select(compound_name=`Metabolite name`, starts_with("180821")) %>%
  pivot_longer(cols=starts_with("180821"), names_to = "filename", values_to="area") %>%
  mutate(polarity="pos")
manual_neg <- read_csv("raw_data/Mesotransect/Area_HILICNeg_EddyTransect.csv", skip = 4) %>%
  filter(`Metabolite name`!="Unknown") %>%
  select(compound_name=`Metabolite name`, starts_with("180821")) %>%
  pivot_longer(cols=starts_with("180821"), names_to = "filename", values_to="area") %>%
  mutate(polarity="neg")
manual_ints <- rbind(manual_pos, manual_neg) %>%
  mutate(compound_name=str_replace_all(compound_name, "Ingalls_|w/o MS2:|Unsettled: ", "")) %>%
  mutate(filename = ifelse(str_detect(filename, "DDA"),
                           str_replace(filename, "Poo", "DDA"),
                           filename)) %>%
  mutate(filename = ifelse(str_detect(filename, "QC"),
                           str_replace(filename, "Poo", "QC"),
                           filename))

# Once again, filter only those that Raafay integrated
clean_ints_transect <- manual_ints %>%
  filter(!str_detect(filename, "DDA")) %>%
  filter(compound_name!="Methyl 3-methylthiopropionate") %>%
  mutate(compound_name=str_replace(compound_name, "Monophosphate, 15N5", "monophosphate, 15N5")) %>%
  left_join(name_conversion, by=c("compound_name"="old")) %>% 
  mutate(compound_name=new) %>% 
  select(-new) %>% 
  left_join(manual_quality, by=c("compound_name", "polarity")) %>%
  filter(integrated=="Y" &in_samps%in%c("Y", "Maybe?", "Kinda?", "Barely", "Kinda")) %>%
  select(-integrated, -in_samps) %>%
  mutate(area=as.numeric(area)) %>%
  anti_join(distinct(., compound_name, polarity) %>% filter(duplicated(compound_name)))

```

```{r transectBMIS}
pooled <- clean_ints_transect %>%
  filter(str_detect(filename, "_Poo_"))

pooled_IS <- pooled %>%
  left_join(clean_stans %>% select(compound_name, compound_type) %>% distinct(), by=c("compound_name")) %>% 
  filter(compound_type=="Internal Standard") %>%
  select(-compound_type, -polarity) %>%
  rename_with(~paste0("IS_", .), -filename) %>%
  add_row(filename=unique(.$filename), IS_compound_name="None", IS_area=1)

initial_CVs <- pooled %>%
  mutate(area=ifelse(str_detect(filename, "Half"), area*2, area)) %>%
  group_by(compound_name) %>%
  summarize(init_CV=cv(area))

standardized_CVs <- full_join(pooled, pooled_IS, by=c("filename")) %>%
  group_by(compound_name, IS_compound_name) %>%
  summarize(norm_CV=cv(area/IS_area), .groups="drop")

chosen_BMIS_transect <- standardized_CVs %>%
  left_join(initial_CVs) %>%
  group_by(compound_name) %>%
  mutate(MIS_improvement=(init_CV - norm_CV)/init_CV) %>% 
  filter(MIS_improvement>min_improvement|IS_compound_name=="None") %>%
  arrange(desc(MIS_improvement)) %>% 
  group_by(compound_name) %>% 
  slice(1) %>%
  select(compound_name, BMIS=IS_compound_name)

BMISed_transect <- clean_ints_transect %>% 
  left_join(chosen_BMIS_transect, by=c("compound_name")) %>%
  left_join(clean_ints_transect %>% rename(IS_area=area) %>% select(-polarity), 
            by=c(BMIS="compound_name", "filename")) %>%
  group_by(compound_name) %>%
  mutate(BMISed_area=area/IS_area*mean(IS_area)) %>%
  mutate(BMISed_area = ifelse(BMIS=="None", area, BMISed_area)) %>%
  mutate(BMISed_area = ifelse(str_detect(filename, "Half")&BMIS=="None", 
                              BMISed_area*2, BMISed_area)) %>%
  select(-IS_area)
```

```{r transectRFs}
transect_rundate <- strptime("180821", "%y%m%d")
transect_stans <- clean_stans %>%
  filter(date_added<=transect_rundate)

rfs_transect <- transect_stans %>%
  left_join(clean_ints_transect, by="compound_name") %>%
  filter(str_detect(filename, "_Std_")) %>% 
  # filter(compound_name=="N6-Acetyl-L-lysine") %>%
  # filter(compound_name=="Dimethylsulfonioacetate") %>%
  # ggplot() + geom_col(aes(x=area,y=filename))
  filter(compound_type!="Internal Standard") %>%
  select(-compound_type) %>%
  mutate(mix=paste0("H2OInMatrix|", paste0(mix, "InMatrix"), "|", paste0(mix, "InH2O"))) %>%
  filter(str_detect(filename, mix)) %>%
  mutate(type=str_replace(str_extract(filename, mix), "Mix[12]", "Mix")) %>%
  select(-mix) %>% 
  group_by(compound_name, type) %>%
  summarise(mean_area=mean(area, na.rm = TRUE), conc_um=unique(conc_um)) %>%
  pivot_wider(names_from = type, values_from=mean_area) %>% 
  mutate(rf_h2o=MixInH2O/conc_um) %>%
  mutate(rf_matrix=(MixInMatrix - H2OInMatrix)/conc_um) %>%
  mutate(rf_ratio=rf_matrix/rf_h2o)

mesotransect_data <- rfs_transect %>% 
  left_join(BMISed_transect, by="compound_name") %>%
  left_join(clean_stans) %>%
  mutate(uM = BMISed_area/rf_h2o * 0.0004/10 * 1/rf_ratio * 2) %>% # 400uL vial, 10L sample, diluted 2x
  mutate(nM=uM*1000) %>%
  select(compound_name, nM, filename, polarity)
```

# Mesocenter things

```{r centerSkylineImport}
manual_pos <- read.csv("raw_data/Mesocenters/Ingalls_Lab_QE_Transition Results_POS.csv") %>%
  mutate(polarity="pos")
manual_neg <- read.csv("raw_data/Mesocenters/Ingalls_Lab_QE_Transition Results_NEG.csv") %>%
  mutate(polarity="neg")
manual_ints <- rbind(manual_pos, manual_neg) %>%
  select(filename=Replicate.Name, compound_name=Precursor.Ion.Name, 
         area=Area, polarity) %>%
  mutate(filename = ifelse(str_detect(filename, "DDA"),
                           str_replace(filename, "Poo", "DDA"),
                           filename)) %>%
  mutate(filename = ifelse(str_detect(filename, "QC"),
                           str_replace(filename, "Poo", "QC"),
                           filename)) %>%
  mutate(compound_name=str_replace(
    compound_name, "Monophosphate, 15N5", "monophosphate, 15N5")) %>%
  left_join(name_conversion, by=c("compound_name"="old")) %>% 
  mutate(compound_name=new) %>% 
  select(-new) %>% 
  left_join(manual_quality, by=c("compound_name", "polarity"))

clean_ints_center <- manual_ints %>%
  filter(integrated=="Y"&in_samps%in%c("Y", "Maybe?", "Kinda?", "Barely", "Kinda")) %>%
  select(-integrated, -in_samps) %>% 
  mutate(area=as.numeric(area))
```

```{r centerBMIS}
pooled <- clean_ints_center %>%
  filter(str_detect(filename, "_Poo_"))

pooled_IS <- pooled %>%
  left_join(clean_stans %>% select(compound_name, compound_type) %>% 
              distinct(), by=c("compound_name")) %>% 
  filter(compound_type=="Internal Standard") %>%
  select(-compound_type, -polarity) %>%
  rename_with(~paste0("IS_", .), -filename) %>%
  add_row(filename=unique(.$filename), IS_compound_name="None", IS_area=1)

initial_CVs <- pooled %>%
  mutate(area=ifelse(str_detect(filename, "Half"), area*2, area)) %>%
  group_by(compound_name) %>%
  summarize(init_CV=cv(area))

standardized_CVs <- full_join(pooled, pooled_IS, by=c("filename")) %>%
  group_by(compound_name, IS_compound_name) %>%
  summarize(norm_CV=cv(area/IS_area), .groups="drop")

chosen_BMIS_center <- standardized_CVs %>%
  left_join(initial_CVs) %>%
  group_by(compound_name) %>%
  mutate(MIS_improvement=(init_CV - norm_CV)/init_CV) %>% 
  filter(MIS_improvement>min_improvement|IS_compound_name=="None") %>%
  arrange(desc(MIS_improvement)) %>%
  group_by(compound_name) %>%
  slice(1) %>%
  select(compound_name, BMIS=IS_compound_name)

BMISed_center <- clean_ints_center %>% 
  left_join(chosen_BMIS_center, by="compound_name") %>%
  left_join(clean_ints_center %>% rename(IS_area=area) %>% select(-polarity), 
            by=c(BMIS="compound_name", "filename")) %>%
  group_by(compound_name) %>%
  mutate(BMISed_area=area/IS_area*mean(IS_area)) %>%
  mutate(BMISed_area = ifelse(BMIS=="None", area, BMISed_area)) %>%
  mutate(BMISed_area = ifelse(str_detect(filename, "Half")&BMIS=="None", 
                              BMISed_area*2, BMISed_area)) %>%
  select(-IS_area)
```

The RFs for the Mesocenter data are a little weird because they weren't run
with H2OinMatrix samples. The good news is that in theory, the RF ratio for
every compound is the same so we can use the RF ratios from Falkor to normalize
these samples instead.

```{r centerRFs}
# Slightly weird because Mesocenter data doesn't have H2OinMatrix samples
# Instead, use rf_ratio from Falkor to swap in
center_rundate <- strptime("180205", "%y%m%d")
center_stans <- clean_stans %>%
  filter(date_added<=center_rundate)

rfs_center <- center_stans %>% 
  left_join(clean_ints_center, by="compound_name") %>% 
  filter(str_detect(filename, "_Std_")) %>%
  filter(compound_type!="Internal Standard") %>%
  select(-compound_type) %>% 
  mutate(mix="InMatrix|InH2O") %>% 
  filter(str_detect(filename, mix)) %>%
  mutate(type=str_extract(filename, mix)) %>%
  select(-mix) %>% 
  group_by(compound_name, type) %>%
  summarize(mean_area=mean(area, na.rm = TRUE), conc_um=unique(conc_um)) %>%
  pivot_wider(names_from = type, values_from=mean_area) %>%
  mutate(rf_h2o=InH2O/conc_um) %>%
  left_join(rfs_falkor %>% select("rf_ratio", "compound_name"), by="compound_name")

mesocenter_data <- rfs_center %>% 
  left_join(BMISed_center, by="compound_name") %>% 
  left_join(clean_stans) %>% 
  mutate(uM = BMISed_area/rf_h2o * 0.0004/10 * 1/rf_ratio * 2) %>% 
  mutate(nM=uM*1000) %>%
  select(compound_name, nM, filename, polarity)
```

# Combine and write out

I'm interested in exporting not only final concentrations, but also the raw
areas and BMIS assignments for debugging purposes. These all go into the
made_data folder.

```{r collect}
## Write out just the raw areas
raw_area_data <- rbind(
  cbind(clean_ints_falkor, cruise="Falkor"),
  cbind(clean_ints_center, cruise="Mesocenter"),
  cbind(clean_ints_transect, cruise="Mesotransect")
) %>%
  left_join(clean_stans, by="compound_name")
write.csv(raw_area_data, file = "made_data/raw_area_data.csv", row.names = FALSE)

## Write out BMISed areas
BMISed_areas <- rbind(
  cbind(BMISed_falkor, cruise="Falkor"),
  cbind(BMISed_center, cruise="Mesocenter"),
  cbind(BMISed_transect, cruise="Mesotransect")
)
write.csv(BMISed_areas, file = "made_data/BMIS_areas.csv", row.names = FALSE)

## Write out RF ratios
rfrs <- rbind(cbind(rfs_falkor, cruise="Falkor"),
              cbind(rfs_transect, cruise="Mesotransect"))
write.csv(rfrs, "made_data/rf_ratios.csv", row.names=FALSE)

## Write out quant data
quant_data <- bind_rows(falkor_data, mesotransect_data, mesocenter_data) %>%
  filter(compound_name!="Choline")
write.csv(quant_data, file = "made_data/longdata.csv", row.names = FALSE)
```

# Perform some quality checks

```{r}
# Fine to have separate H2OinMatrix cols, different filenames between each
rfrs <- read_csv("made_data/rf_ratios.csv") %>%
  mutate(H2OinMatrix=ifelse(is.na(H2OinMatrix), H2OInMatrix, H2OinMatrix)) %>%
  select(-H2OInMatrix) %>%
  select(compound_name, rf_ratio, cruise) %>%
  pivot_wider(names_from = cruise, values_from = rf_ratio)

with(rfrs, cor(Falkor, Mesotransect, use="pair"))

ggplot(rfrs, aes(x=Falkor, y=Mesotransect, label=compound_name)) +
  geom_point()
plotly::ggplotly()
rfrs %>%
  filter(compound_name!="Choline") %>%
  ggplot(aes(x=Falkor, y=Mesotransect, label=compound_name)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) +
  lims(x=c(0, NA), y=c(0, NA)) +
  coord_fixed()
```