### 2023-08-29 targeted/targeted_control.Rmd
Complete rerun from MW file download

---
title: "Peak area control doc"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, purl = TRUE)
options("readr.num_columns" = 0)

library(tidyverse)

already_good <- 0.1
min_improvement <- 0.4

cv <- function(x)sd(x)/mean(x)

clean_stans <- read.csv("../metadata/clean_stans.csv")
name_conversion <- clean_stans %>%
  select(new=compound_name, old=name_old) %>%
  distinct()
date_filter <- clean_stans %>%
  select(compound_name=name_old, date_added) %>%
  distinct()
file_metadata <- read.csv("../metadata/filled_file_metadata.csv")
```

# Convert raw data ----

First step is getting the data out of Skyline/MSDIAL. This is doable by script
if the original .sky and .mdf files are available, as they should be in this
case in the raw_data folder broken down by cruise.

We also need the console interfaces to both Skyline and MSDIAL.
MSDIAL's comes with the app and is pointed to with a raw file path to wherever
the MSDIAL files were extracted, and Skyline requires a separate download of
their command line version.

Mesotransect was integrated with MSDIAL, Falkor and Mesocenter used Skyline.

```{r grabRAnotes}
if(!file.exists("raw_data/Raafay_HILIC_notes.csv")){
  library(googlesheets4)
  if(!gs4_has_token())stop("Refresh Google token")
  # googlesheets4::gs4_auth()

  manual_pos <- read_sheet(ss = "1lU7g5rfem81s_0fqMP93yzegogCN29JJEdsIESfAX7c", 
                           skip = 1, sheet = "HILIC_POS") %>%
    select(compound_name=`Compound name`,
           integrated=`Integrated? (Y/N)`,
           in_samps=`In samples? (Y/N)`) %>%
    mutate(polarity="pos")
  manual_neg <- read_sheet(ss = "1lU7g5rfem81s_0fqMP93yzegogCN29JJEdsIESfAX7c",
                           sheet = "HILIC_NEG") %>%
    select(compound_name=`Compound name`,
           integrated=`Integrated? (Y/N)`,
           in_samps=`In samples? (Y/N)`) %>%
    mutate(polarity="neg")
  manual_quality <- rbind(manual_pos, manual_neg) %>%
    mutate(compound_name=str_replace(compound_name, "_ ", ", ")) %>%
    filter(!is.na(compound_name)) %>%
    filter(compound_name!="Internal Standards") %>%
  left_join(name_conversion, by=c(compound_name="old")) %>%
  mutate(compound_name=new) %>%
  select(-new)
  write.csv(manual_quality, file = "raw_data/Raafay_HILIC_notes.csv", 
            row.names = FALSE)
}
manual_quality <- read.csv("raw_data/Raafay_HILIC_notes.csv") 
```

SkylineRunner.exe output relating to exporting files:
Exporting reports
ÚÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÂÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄ¿
³--report-name=<name>          ³The name of a report to export as it appears ³
³                              ³in the Skyline Export Report form.           ³
ÃÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÅÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄ´
³--report-file=path/to/file.csv³The path to export the report to. Required if³
³                              ³--report-name is specified.                  ³
ÃÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÅÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄ´
³--report-add=path/to/file.skyr³Add reports from a saved report template file³
³                              ³to the running program.                      ³
ÃÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÅÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄ´
³--report-conflict-resolution= ³Specifies how to resolve report name         ³
³<overwrite | skip>            ³conflicts, by either overwriting or skipping ³
³                              ³them, when using --report-add (default is to ³
³                              ³output an error message for conflicts)       ³
ÃÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÅÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄÄ´
³--report-format=<csv | tsv>   ³CSV for comma-separated reports (or semicolon³
³                              ³separated, depending on your localization) or³
³                              ³TSV for tab separated reports [default CSV]  ³

```{r Falkor Skyline data extraction}
FK_POS_cmd <- paste(
  "SkylineRunner.exe",
  "--in=raw_data/Falkor/HILIC_QE_POSITIVE_Falkor_1.sky", 
  '--report-name="Ingalls_Lab_QE_Transition Results"',
  '--report-file="raw_data/Falkor/Ingalls_Lab_QE_Transition Results_POS.csv"',
  '--report-format="csv"'
)
system(FK_POS_cmd)

FK_neg_cmd <- paste(
  "SkylineRunner.exe",
  "--in=raw_data/Falkor/HILIC_QE_NEGATIVE_Falkor.sky", 
  '--report-name="Ingalls_Lab_QE_Transition Results"',
  '--report-file="raw_data/Falkor/Ingalls_Lab_QE_Transition Results_NEG.csv"',
  '--report-format="csv"'
)
system(FK_neg_cmd)
```

```{r Mesocenter Skyline data extraction}
MC_POS_cmd <- paste(
  "SkylineRunner.exe",
  "--in=raw_data/Mesocenters/HILIC_QE_POSITIVE_Wei_HRM_DepthProfile.sky", 
  '--report-name="Ingalls_Lab_QE_Transition Results"',
  '--report-file="raw_data/Mesocenters/Ingalls_Lab_QE_Transition Results_POS.csv"',
  '--report-format="csv"'
)
system(MC_POS_cmd)

FK_NEG_cmd <- paste(
  "SkylineRunner.exe",
  "--in=raw_data/Mesocenters/HILIC_QE_NEGATIVE_Wei_HRM_DepthProfile.sky", 
  '--report-name="Ingalls_Lab_QE_Transition Results"',
  '--report-file="raw_data/Mesocenters/Ingalls_Lab_QE_Transition Results_NEG.csv"',
  '--report-format="csv"'
)
system(FK_NEG_cmd)
```

```{r Mesotransect MSDIAL autoexport things not yet}
# Not implemented yet, check on thread below for updates
# http://www.metabolomics-forum.com/index.php?topic=1820.0
# Instead, open an existing .mtd2 file, click the "Export" tab, select "Alignment result", specify the output directory as the current working directory, check the Raw Data matrix (Area) box , keep the output format as msp, and hit the "Export" button at the bottom.
```

# Falkor things

Start with the Falkor data, since it's nice and neat. Here we're pulling from
the raw_data/Falkor folder, which has the Skyline files, their output, and
will soon have Raafay's notes on peak quality.

These CSV files were obtained from Skyline by exporting the transition results,
with a "_POS" or "_NEG" added to distinguish them.

```{r falkorSkylineImport}
# Read in the data and clean it up a little bit
manual_pos <- read.csv("raw_data/Falkor/Ingalls_Lab_QE_Transition Results_POS.csv") %>%
  mutate(polarity="pos")
manual_neg <- read.csv("raw_data/Falkor/Ingalls_Lab_QE_Transition Results_NEG.csv") %>%
  mutate(polarity="neg")
manual_ints <- rbind(manual_pos, manual_neg) %>%
  select(filename=Replicate.Name, compound_name=Precursor.Ion.Name, 
         area=Area, polarity) %>% 
  mutate(filename = ifelse(str_detect(filename, "DDA"),
                           str_replace(filename, "Poo", "DDA"),
                           filename)) %>%
  mutate(filename = ifelse(str_detect(filename, "QC"),
                           str_replace(filename, "Poo", "QC"),
                           filename)) %>% 
  mutate(compound_name=str_replace(
    compound_name, "Monophosphate, 15N5", "monophosphate, 15N5")) %>%
  left_join(name_conversion, by=c("compound_name"="old")) %>% 
  mutate(compound_name=new) %>% 
  select(-new) %>% 
  left_join(manual_quality, by=c("compound_name", "polarity"))

# Filter for only those peaks that Raafay actually integrated
clean_ints_falkor <- manual_ints %>%
  filter(integrated=="Y"&in_samps%in%c("Y", "Maybe?", "Kinda?", "Barely", "Kinda")) %>%
  select(-integrated, -in_samps) %>% 
  mutate(area=as.numeric(area))
```

```{r falkorBMIS}
pooled <- clean_ints_falkor %>%
  filter(str_detect(filename, "_Poo_"))

pooled_IS <- pooled %>%
  left_join(clean_stans %>% select(compound_name, compound_type) %>% distinct(), by=c("compound_name")) %>% 
  filter(compound_type=="Internal Standard") %>%
  select(-compound_type, -polarity) %>%
  rename_with(~paste0("IS_", .), -filename) %>%
  add_row(filename=unique(.$filename), IS_compound_name="None", IS_area=1)

initial_CVs <- pooled %>%
  mutate(area=ifelse(str_detect(filename, "Half"), area*2, area)) %>%
  group_by(compound_name) %>%
  summarize(init_CV=cv(area))

standardized_CVs <- full_join(pooled, pooled_IS, by=c("filename")) %>%
  group_by(compound_name, IS_compound_name) %>%
  summarize(norm_CV=cv(area/IS_area), .groups="drop")

chosen_BMIS_falkor <- standardized_CVs %>%
  left_join(initial_CVs) %>%
  group_by(compound_name) %>%
  mutate(MIS_improvement=(init_CV - norm_CV)/init_CV) %>% 
  filter(MIS_improvement>min_improvement|IS_compound_name=="None") %>%
  arrange(desc(MIS_improvement)) %>%
  group_by(compound_name) %>%
  slice(1) %>%
  select(compound_name, BMIS=IS_compound_name)

BMISed_falkor <- clean_ints_falkor %>%
  left_join(chosen_BMIS_falkor, by="compound_name") %>%
  left_join(clean_ints_falkor %>% rename(IS_area=area) %>% select(-polarity), 
            by=c(BMIS="compound_name", "filename")) %>%
  group_by(compound_name) %>%
  mutate(BMISed_area=area/IS_area*mean(IS_area)) %>%
  mutate(BMISed_area = ifelse(BMIS=="None", area, BMISed_area)) %>%
  mutate(BMISed_area = ifelse(str_detect(filename, "Half")&BMIS=="None", 
                              BMISed_area*2, BMISed_area)) %>%
  select(-IS_area)
```

```{r falkorRFs}
falkor_rundate <- 190715
falkor_stans <- clean_stans %>%
  filter(date_added<=falkor_rundate)

rfs_falkor <- falkor_stans %>%
  left_join(BMISed_falkor, by=c("compound_name", "polarity")) %>%
  filter(str_detect(filename, "_Std_")) %>%
  filter(compound_type!="Internal Standard") %>%
  select(-compound_type) %>%
  mutate(mix=paste0("H2OinMatrix|", paste0(mix, "InMatrix"), "|", paste0(mix, "InH2O"))) %>%
  filter(str_detect(filename, mix)) %>%
  mutate(type=str_replace(str_extract(filename, mix), "Mix[12]", "Mix")) %>%
  select(-mix) %>%
  group_by(compound_name, type) %>%
  summarize(mean_area=mean(area, na.rm = TRUE), conc_um=unique(conc_um)) %>%
  pivot_wider(names_from = type, values_from=mean_area) %>%
  mutate(rf_h2o=MixInH2O/conc_um) %>%
  mutate(rf_matrix=(MixInMatrix - H2OinMatrix)/conc_um) %>%
  mutate(rf_ratio=rf_matrix/rf_h2o)

# 400uL vial, 10L sample, diluted x2
falkor_data <- rfs_falkor %>% 
  left_join(BMISed_falkor, by="compound_name") %>%
  left_join(clean_stans) %>%
  mutate(uM = BMISed_area/rf_h2o * 0.0004/10 * 1/rf_ratio * 2) %>% 
  mutate(nM=uM*1000) %>%
  select(compound_name, nM, filename, polarity)
```

# Mesotransect things

Continue with the MESOSCOPE transect data. Here again, we use Raafay's notes
on peak quality to remove the gross peaks.

```{r transectMsdialImport}

manual_neg <- read_delim("raw_data/Mesotransect/MSDial_neg_data/Area_0_20235251247.txt", skip = 4, delim = "\t") %>%
  filter(`Metabolite name`!="Unknown") %>%
  select(compound_name=`Metabolite name`, starts_with("180821")) %>%
  pivot_longer(cols=starts_with("180821"), names_to = "filename", values_to="area") %>%
  mutate(polarity="neg")
manual_pos <- read_delim("raw_data/Mesotransect/MSDial_pos_data/Area_2_2023525133.txt", skip = 4, delim = "\t") %>%
  filter(`Metabolite name`!="Unknown") %>%
  select(compound_name=`Metabolite name`, starts_with("180821")) %>%
  pivot_longer(cols=starts_with("180821"), names_to = "filename", values_to="area") %>%
  mutate(polarity="pos")
manual_ints <- rbind(manual_pos, manual_neg) %>%
  mutate(compound_name=str_replace_all(compound_name, "Ingalls_|w/o MS2:|Unsettled: ", "")) %>%
  mutate(filename = ifelse(str_detect(filename, "DDA"),
                           str_replace(filename, "Poo", "DDA"),
                           filename)) %>%
  mutate(filename = ifelse(str_detect(filename, "QC"),
                           str_replace(filename, "Poo", "QC"),
                           filename)) %>%
  anti_join(date_filter %>% filter(date_added>180821))

# Once again, filter only those that Raafay integrated
clean_ints_transect <- manual_ints %>%
  filter(!str_detect(filename, "DDA")) %>%
  filter(compound_name!="Methyl 3-methylthiopropionate") %>%
  filter(compound_name!="N(e)-Acetyl-Lysine") %>%
  filter(compound_name!="Valine") %>%
  mutate(compound_name=str_replace(compound_name, "Monophosphate, 15N5", "monophosphate, 15N5")) %>%
  left_join(name_conversion, by=c("compound_name"="old")) %>% 
  mutate(compound_name=ifelse(is.na(new), compound_name, new)) %>% 
  select(-new) %>% 
  left_join(manual_quality, by=c("compound_name", "polarity")) %>%
  filter(integrated=="Y" &in_samps%in%c("Y", "Maybe?", "Kinda?", "Barely", "Kinda")) %>%
  select(-integrated, -in_samps) %>%
  mutate(area=as.numeric(area)) %>%
  anti_join(distinct(., compound_name, polarity) %>% filter(duplicated(compound_name)))

# Fix sample 180821_Smp_MS6C315m_C which didn't get extraction standards added
ext_stans_list <- c(
  "Isethionic acid, 13C2",
  "L-Cysteic acid, 2H3",
  "Sulfoacetic acid, 13C2",
  "3-Sulfolactate, 13C3",
  "Taurine, 2H4"
)
update_df <- clean_ints_transect %>%
  filter(str_detect(filename, "Smp")) %>%
  filter(polarity=="neg") %>%
  filter(compound_name%in%ext_stans_list) %>%
  group_by(compound_name) %>%
  summarise(area=median(area)) %>%
  mutate(filename="180821_Smp_MS6C315m_C")
clean_ints_transect <- clean_ints_transect %>%
  rows_update(update_df, by = c("compound_name", "filename"))
```

```{r transectBMIS}
pooled <- clean_ints_transect %>%
  filter(str_detect(filename, "_Poo_"))

pooled_IS <- pooled %>%
  left_join(clean_stans %>% select(compound_name, compound_type) %>% distinct(), by=c("compound_name")) %>% 
  filter(compound_type=="Internal Standard") %>%
  select(-compound_type, -polarity) %>%
  rename_with(~paste0("IS_", .), -filename) %>%
  add_row(filename=unique(.$filename), IS_compound_name="None", IS_area=1)

initial_CVs <- pooled %>%
  mutate(area=ifelse(str_detect(filename, "Half"), area*2, area)) %>%
  group_by(compound_name) %>%
  summarize(init_CV=cv(area))

standardized_CVs <- full_join(pooled, pooled_IS, by=c("filename")) %>%
  group_by(compound_name, IS_compound_name) %>%
  summarize(norm_CV=cv(area/IS_area), .groups="drop")

chosen_BMIS_transect <- standardized_CVs %>%
  left_join(initial_CVs) %>%
  group_by(compound_name) %>%
  mutate(MIS_improvement=(init_CV - norm_CV)/init_CV) %>% 
  filter(MIS_improvement>min_improvement|IS_compound_name=="None") %>%
  arrange(desc(MIS_improvement)) %>% 
  group_by(compound_name) %>% 
  slice(1) %>%
  select(compound_name, BMIS=IS_compound_name)

BMISed_transect <- clean_ints_transect %>% 
  left_join(chosen_BMIS_transect, by=c("compound_name")) %>%
  left_join(clean_ints_transect %>% rename(IS_area=area) %>% select(-polarity), 
            by=c(BMIS="compound_name", "filename")) %>%
  group_by(compound_name) %>%
  mutate(BMISed_area=area/IS_area*mean(IS_area)) %>%
  mutate(BMISed_area = ifelse(BMIS=="None", area, BMISed_area)) %>%
  mutate(BMISed_area = ifelse(str_detect(filename, "Half")&BMIS=="None", 
                              BMISed_area*2, BMISed_area)) %>%
  select(-IS_area)
```

```{r transectRFs}
transect_rundate <- strptime("180821", "%y%m%d")
transect_stans <- clean_stans %>%
  filter(strptime(date_added, "%y%m%d")<=transect_rundate)

rfs_transect <- transect_stans %>%
  left_join(clean_ints_transect, by=c("compound_name", "polarity")) %>%
  filter(str_detect(filename, "_Std_")) %>% 
  # filter(compound_name=="N6-Acetyl-L-lysine") %>%
  # filter(compound_name=="Dimethylsulfonioacetate") %>%
  # ggplot() + geom_col(aes(x=area,y=filename))
  filter(compound_type!="Internal Standard") %>%
  select(-compound_type) %>%
  mutate(mix=paste0("H2OInMatrix|", paste0(mix, "InMatrix"), "|", paste0(mix, "InH2O"))) %>%
  filter(str_detect(filename, mix)) %>%
  mutate(type=str_replace(str_extract(filename, mix), "Mix[12]", "Mix")) %>%
  select(-mix) %>% 
  group_by(compound_name, type) %>%
  summarise(mean_area=mean(area, na.rm = TRUE), conc_um=unique(conc_um)) %>%
  pivot_wider(names_from = type, values_from=mean_area) %>% 
  mutate(rf_h2o=MixInH2O/conc_um) %>%
  mutate(rf_matrix=(MixInMatrix - H2OInMatrix)/conc_um) %>%
  mutate(rf_ratio=rf_matrix/rf_h2o)

mesotransect_data <- rfs_transect %>% 
  left_join(BMISed_transect, by="compound_name", multiple="all") %>%
  left_join(clean_stans) %>%
  mutate(uM = BMISed_area/rf_h2o * 0.0004/10 * 1/rf_ratio * 2) %>% # 400uL vial, 10L sample, diluted 2x
  mutate(nM=uM*1000) %>%
  select(compound_name, nM, filename, polarity)
```

```{r transect fix sucrose/trehalose integrations, eval=FALSE}
# Sucrose and trehalose got integrated together which is obnoxious because we 
# have IS for each of those so it'd be really nice to get those good numbers
# Re-integrating using RaMS

library(RaMS)
# msdata <- list.files("../tmzMLs/neg", full.names = TRUE, 
#                      pattern = "170706|180821|180205") %>% 
#   grabMSdata(verbosity = 0)
# tresuc_msdata <- msdata$MS1[mz%between%pmppm(341.10839, 10)]
# IS_msdata <- msdata$MS1[mz%between%pmppm(353.1497, 10)]

suc_data <- tresuc_msdata[rt%between%c(11.2, 12.1)] %>%
  group_by(filename) %>%
  summarize(area=sum(int))
suc_IS_data <- IS_msdata[rt%between%c(11.2, 12.1)] %>%
  group_by(filename) %>%
  summarize(IS_area=sum(int))
# ggplot(suc_IS_data) + geom_col(aes(x=filename, y=IS_area))
suc_nM <- left_join(suc_data, suc_IS_data) %>%
  mutate(nM=area/IS_area*5000 * 0.0004/10 * 2) %>%
  mutate(compound_name="Sucrose")

tre_data <- tresuc_msdata[rt%between%c(12.1, 12.9)] %>%
  group_by(filename) %>%
  summarize(area=sum(int))
tre_IS_data <- IS_msdata[rt%between%c(12.1, 12.9)] %>%
  group_by(filename) %>%
  summarize(IS_area=sum(int))
# ggplot(suc_IS_data) + geom_col(aes(x=filename, y=IS_area))
tre_nM <- left_join(tre_data, tre_IS_data) %>%
  mutate(nM=area/IS_area*5000 * 0.0004/10 * 2) %>%
  mutate(compound_name="Trehalose")

update_df <- bind_rows(tre_nM, suc_nM) %>%
  select(compound_name, nM, filename) %>%
  mutate(polarity="neg") %>%
  mutate(filename=str_remove(filename, ".tmzML"))
mesotransect_data <- mesotransect_data %>%
  rows_upsert(update_df, by=c("compound_name", "nM", "filename", "polarity"))
```

# Mesocenter things

```{r centerSkylineImport}
manual_pos <- read.csv("raw_data/Mesocenters/Ingalls_Lab_QE_Transition Results_POS.csv") %>%
  mutate(polarity="pos")
manual_neg <- read.csv("raw_data/Mesocenters/Ingalls_Lab_QE_Transition Results_NEG.csv") %>%
  mutate(polarity="neg")
manual_ints <- rbind(manual_pos, manual_neg) %>%
  select(filename=Replicate.Name, compound_name=Precursor.Ion.Name, 
         area=Area, polarity) %>%
  mutate(filename = ifelse(str_detect(filename, "DDA"),
                           str_replace(filename, "Poo", "DDA"),
                           filename)) %>%
  mutate(filename = ifelse(str_detect(filename, "QC"),
                           str_replace(filename, "Poo", "QC"),
                           filename)) %>%
  mutate(compound_name=str_replace(
    compound_name, "Monophosphate, 15N5", "monophosphate, 15N5")) %>%
  left_join(name_conversion, by=c("compound_name"="old")) %>% 
  mutate(compound_name=ifelse(is.na(new), compound_name, new)) %>% 
  select(-new) %>% 
  left_join(manual_quality, by=c("compound_name", "polarity"))

clean_ints_center <- manual_ints %>%
  filter(integrated=="Y"&in_samps%in%c("Y", "Maybe?", "Kinda?", "Barely", "Kinda")) %>%
  select(-integrated, -in_samps) %>% 
  mutate(area=as.numeric(area)) %>%
  filter(!compound_name%in%c("Turicine", "Hydroxyisoleucine"))
```

```{r centerBMIS}
pooled <- clean_ints_center %>%
  filter(str_detect(filename, "_Poo_"))

pooled_IS <- pooled %>%
  left_join(clean_stans %>% select(compound_name, compound_type) %>% 
              distinct(), by=c("compound_name")) %>% 
  filter(compound_type=="Internal Standard") %>%
  select(-compound_type, -polarity) %>%
  rename_with(~paste0("IS_", .), -filename) %>%
  add_row(filename=unique(.$filename), IS_compound_name="None", IS_area=1)

initial_CVs <- pooled %>%
  mutate(area=ifelse(str_detect(filename, "Half"), area*2, area)) %>%
  group_by(compound_name) %>%
  summarize(init_CV=cv(area))

standardized_CVs <- full_join(pooled, pooled_IS, by=c("filename")) %>%
  group_by(compound_name, IS_compound_name) %>%
  summarize(norm_CV=cv(area/IS_area), .groups="drop")

chosen_BMIS_center <- standardized_CVs %>%
  left_join(initial_CVs) %>%
  group_by(compound_name) %>%
  mutate(MIS_improvement=(init_CV - norm_CV)/init_CV) %>% 
  filter(MIS_improvement>min_improvement|IS_compound_name=="None") %>%
  arrange(desc(MIS_improvement)) %>%
  group_by(compound_name) %>%
  slice(1) %>%
  select(compound_name, BMIS=IS_compound_name)

BMISed_center <- clean_ints_center %>% 
  left_join(chosen_BMIS_center, by="compound_name") %>%
  left_join(clean_ints_center %>% rename(IS_area=area) %>% select(-polarity), 
            by=c(BMIS="compound_name", "filename")) %>%
  group_by(compound_name) %>%
  mutate(BMISed_area=area/IS_area*mean(IS_area)) %>%
  mutate(BMISed_area = ifelse(BMIS=="None", area, BMISed_area)) %>%
  mutate(BMISed_area = ifelse(str_detect(filename, "Half")&BMIS=="None", 
                              BMISed_area*2, BMISed_area)) %>%
  select(-IS_area)
```

The RFs for the Mesocenter data are a little weird because they weren't run
with H2OinMatrix samples. The good news is that in theory, the RF ratio for
every compound is the same so we can use the RF ratios from Falkor to normalize
these samples instead.

```{r centerRFs}
# Slightly weird because Mesocenter data doesn't have H2OinMatrix samples
center_rundate <- strptime("180205", "%y%m%d")
center_stans <- clean_stans %>%
  filter(strptime(date_added, "%y%m%d")<=center_rundate)

custom_h2o_mat <- clean_ints_center %>%
  filter(str_detect(filename, "Poo_Full")) %>%
  group_by(compound_name) %>%
  summarise(newH2OInMatrix=mean(area)*(1-(62.94*100/80)/100))

rfs_center <- center_stans %>% 
  left_join(clean_ints_center, by=c("compound_name", "polarity")) %>%
  filter(str_detect(filename, "_Std_")) %>%
  filter(compound_type!="Internal Standard") %>%
  select(-compound_type) %>% 
  mutate(mix="InMatrix|InH2O") %>% 
  filter(str_detect(filename, mix)) %>%
  mutate(type=str_extract(filename, mix)) %>%
  select(-mix) %>% 
  group_by(compound_name, type) %>%
  summarize(mean_area=mean(area, na.rm = TRUE), conc_um=unique(conc_um)) %>%
  pivot_wider(names_from = type, values_from=mean_area) %>%
  left_join(custom_h2o_mat) %>%
  # left_join(rfs_falkor) %>%
  mutate(rf_h2o=InH2O/conc_um) %>%
  mutate(rf_matrix=(InMatrix - newH2OInMatrix)/conc_um) %>%
  mutate(rf_ratio=rf_matrix/rf_h2o)

mesocenter_data <- rfs_center %>% 
  left_join(BMISed_center, by="compound_name") %>% 
  left_join(clean_stans) %>% 
  mutate(uM = BMISed_area/rf_h2o * 0.0004/10 * 1/rf_ratio * 2) %>%
  mutate(nM=uM*1000) %>%
  select(compound_name, nM, filename, polarity)
```

# Combine and write out

I'm interested in exporting not only final concentrations, but also the raw
areas and BMIS assignments for debugging purposes. These all go into the
made_data folder.

```{r collect}
## Write out just the raw areas
raw_area_data <- rbind(
  cbind(clean_ints_falkor, cruise="Falkor"),
  cbind(clean_ints_center, cruise="Mesocenter"),
  cbind(clean_ints_transect, cruise="Mesotransect")
) %>%
  left_join(clean_stans, by=c("compound_name", "polarity"))
write.csv(raw_area_data, file = "made_data/raw_area_data.csv", row.names = FALSE)

## Write out BMISed areas
BMISed_areas <- rbind(
  cbind(BMISed_falkor, cruise="Falkor"),
  cbind(BMISed_center, cruise="Mesocenter"),
  cbind(BMISed_transect, cruise="Mesotransect")
)
write.csv(BMISed_areas, file = "made_data/BMIS_areas.csv", row.names = FALSE)

## Write out RF ratios
rfrs <- rbind(cbind(rfs_falkor, cruise="Falkor"),
              cbind(rfs_transect, cruise="Mesotransect"))
write.csv(rfrs, "made_data/rf_ratios.csv", row.names=FALSE)

## Write out quant data
quant_data <- bind_rows(falkor_data, mesotransect_data, mesocenter_data) %>%
  filter(compound_name!="Choline") %>%
  mutate(nM=ifelse(compound_name=="Homarine", nM/25, nM))
write.csv(quant_data, file = "made_data/longdata.csv", row.names = FALSE)
write_csv(quant_data, file = "all_peaks.csv")
```

```{r write MESOSCOPE to Excel}
library(openxlsx)
wb <- createWorkbook(creator = "William Kumler, wkumler@uw.edu")
addWorksheet(wb, sheetName = "Sample data")
bind_rows(mesotransect_data, mesocenter_data) %>%
  pivot_wider(names_from = filename, values_from = nM) %>%
  writeData(wb = wb, sheet = "Sample data")
freezePane(wb, sheet = "Sample data", firstRow = TRUE, firstCol = TRUE)
addWorksheet(wb, "File metadata")
file_metadata %>%
  filter(cruise%in%c("MT", "MC")) %>%
  mutate(filename=str_remove(filename, ".mzML")) %>%
  distinct(filename, samp_type, station, cast, depth, abs_depth, rosette,date_run) %>%
  rename(sample_name=filename) %>%
  writeData(wb = wb, sheet = "File metadata")
addWorksheet(wb = wb, sheetName = "Compound metadata")
clean_stans %>%
  select(compound_name, compound_type, formula, stan_rt, stan_mz, 
         kegg_id, polarity, peak_quality) %>%
  writeData(wb = wb, sheet = "Compound metadata")

addWorksheet(wb=wb, sheetName = "README")
data.frame(`README for Ingalls MESOSCOPE data`=c(
  "This file was produced by William Kumler in the Ingalls lab at the University of Washington",
  "This file contains concentrations of compounds in seawater measured via HPLC-MS",
  "in a targeted (selected subset of known compounds) method.",
  "",
  "The first sheet (Sample data) contains the concentrations of various molecules in each sample",
  "taken during the MESOSCOPE cruise in 2017. These values are in nanomoles per",
  "liter of seawater (nM) from the particulate phase (particles collected onto",
  "0.2um Durapore filters).",
  "Missing values are shown as empty cells and infinite values (usually in the standards) are shown as #NUM!",
  "",
  "The second sheet (File metadata) contains information about the samples themselves, including",
  "the depth, time, and station from which they were collected. ",
  "sample_name = Name of the sample as shown in the Sample data Excel sheet",
  "samp_type = Type of sample, whether biological sample (Smp), blank (Blk), pooled (Poo), or standard (Std)",
  "station = Station from which the sample was collected, 1-15 plus L1 and L2",
  "cast = Which cast the sample was taken from",
  "depth = Depth descriptor of where in the water column the sample was collected",
  "abs_depth = Absolute depth. Accounts for differences from DCM and is determined from the CTD itself",
  "rosette = Which Niskin was this sample taken from? Largely incomplete",
  "date_run = The date on which the HPLC-MS run was started",
  "",
  "The third sheet (Compound metadata) contains information about the compounds measured",
  "during this analysis. Not all compounds were detected and passed quality control",
  "so there are more compounds here than found in the Sample data sheet.",
  "compound_name = Name of the compound, as found in the Sample data sheet",
  "compound_type = Broad compound class as determined rather arbitrarily by myself",
  "formula = Molecular formula of the compound. Isotopes are labeled in parentheses",
  "stan_rt = Estimated retention time of the compound, in minutes",
  "stan_mz = Mass-to-charge ratio of the molecule at which it's detected in the mass spec",
  "kegg_id = KEGG (Kyoto Encyclopedia of Genes and Genomes) ID: https://www.genome.jp/kegg/",
  "polarity = The polarity at which the molecule was detected in the mass spec, either 'pos' for positive mode or 'neg' for negative mode",
  "peak_quality = The typical quality of the chromatographic peak. Yes is good, No is bad, and see the Ingalls Standards repo for more details on Problematic compounds",
  "",
  paste("File created on", Sys.Date())
)) %>%
  writeData(wb = wb, sheet = "README")
saveWorkbook(wb, file = "MESOSCOPE_Targeted_nM.xlsx", overwrite = TRUE)
```

```{r write Falkor to Excel}
library(openxlsx)
wb <- createWorkbook(creator = "William Kumler, wkumler@uw.edu")
addWorksheet(wb, sheetName = "Sample data")
falkor_data %>%
  pivot_wider(names_from = filename, values_from = nM) %>%
  writeData(wb = wb, sheet = "Sample data")
freezePane(wb, sheet = "Sample data", firstRow = TRUE, firstCol = TRUE)
addWorksheet(wb, "File metadata")
file_metadata %>%
  filter(cruise=="FK") %>%
  mutate(filename=str_remove(filename, ".mzML")) %>%
  distinct(filename, samp_type, station, cast, depth, abs_depth, rosette, date_run) %>%
  rename(sample_name=filename) %>%
  writeData(wb = wb, sheet = "File metadata")
addWorksheet(wb = wb, sheetName = "Compound metadata")
clean_stans %>%
  select(compound_name, compound_type, formula, stan_rt, stan_mz, 
         kegg_id, polarity, peak_quality) %>%
  writeData(wb = wb, sheet = "Compound metadata")

addWorksheet(wb=wb, sheetName = "README")
data.frame(`README for Ingalls Falkor data`=c(
  "This file was produced by William Kumler in the Ingalls lab at the University of Washington",
  "This file contains concentrations of compounds in seawater measured via HPLC-MS",
  "in a targeted (selected subset of known compounds, manually integrated) method.",
  "",
  "The first sheet (Sample data) contains the concentrations of various molecules in each sample",
  "taken during the Falkor cruise in 2018. These values are in nanomoles per",
  "liter of seawater (nM) from the particulate phase (particles collected onto",
  "0.2um Durapore filters).",
  "Missing values are shown as empty cells and infinite values (usually in the standards) are shown as #NUM!",
  "",
  "The second sheet (File metadata) contains information about the samples themselves, including",
  "the depth, time, and station from which they were collected. ",
  "sample_name = Name of the sample as shown in the Sample data Excel sheet",
  "samp_type = Type of sample, whether biological sample (Smp), blank (Blk), pooled (Poo), or standard (Std)",
  "station = Station from which the sample was collected",
  "cast = Which cast the sample was taken from",
  "depth = Depth descriptor of where in the water column the sample was collected",
  "abs_depth = Absolute depth. Accounts for differences from DCM and is determined from the CTD itself",
  "rosette = Which Niskin was this sample taken from? Largely incomplete",
  "date_run = The date on which the HPLC-MS run was started",
  "",
  "The third sheet (Compound metadata) contains information about the compounds measured",
  "during this analysis. Not all compounds were detected and passed quality control",
  "so there are more compounds here than found in the Sample data sheet.",
  "compound_name = Name of the compound, as found in the Sample data sheet",
  "compound_type = Broad compound class as determined rather arbitrarily by myself",
  "formula = Molecular formula of the compound. Isotopes are labeled in parentheses",
  "stan_rt = Estimated retention time of the compound, in minutes",
  "stan_mz = Mass-to-charge ratio of the molecule at which it's detected in the mass spec",
  "kegg_id = KEGG (Kyoto Encyclopedia of Genes and Genomes) ID: https://www.genome.jp/kegg/",
  "polarity = The polarity at which the molecule was detected in the mass spec, either 'pos' for positive mode or 'neg' for negative mode",
  "peak_quality = The typical quality of the chromatographic peak. Yes is good, No is bad, and see the Ingalls Standards repo for more details on Problematic compounds",
  "",
  paste("File created on", Sys.Date())
)) %>%
  writeData(wb = wb, sheet = "README")
saveWorkbook(wb, file = "Falkor_Targeted_nM.xlsx", overwrite = TRUE)
```

# Perform some quality checks

```{r}
# Fine to have separate H2OinMatrix cols, different filenames between each
rfrs <- read_csv("made_data/rf_ratios.csv") %>%
  mutate(H2OinMatrix=ifelse(is.na(H2OinMatrix), H2OInMatrix, H2OinMatrix)) %>%
  select(-H2OInMatrix) %>%
  select(compound_name, rf_ratio, cruise) %>%
  pivot_wider(names_from = cruise, values_from = rf_ratio)

with(rfrs, cor(Falkor, Mesotransect, use="pair"))

ggplot(rfrs, aes(x=Falkor, y=Mesotransect, label=compound_name)) +
  geom_point()
plotly::ggplotly()
rfrs %>%
  filter(compound_name!="Choline") %>%
  ggplot(aes(x=Falkor, y=Mesotransect, label=compound_name)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) +
  lims(x=c(0, NA), y=c(0, NA)) +
  coord_fixed()
```