### 2024-11-15 manuscript/response_to_reviewers.txt
Fixing line numbers and tidying response

> Reviewer 1 | 04 Sep 2024 | 01:18
> #1
> This study examines the impact of mesoscale eddies on the metabolome and composition 
> of particulate organic matter in the North Pacific Subtropical Gyre. It fills a gap 
> in previous research by exploring the molecular composition of particulate organic 
> matter. The use both of targeted and untargeted metabolomics mass spectrometry to 
> detect molecules sensitive to environmental perturbation. The study finds that many 
> metabolites track biomass trends. Metabolites linked to eukaryotic phytoplankton are 
> enriched at the deep chlorophyll maximum of cyclonic features, while degradation 
> products are enriched in anticyclones. Additionally, depth variability and vertical 
> displacement of isopycnal layer constrain the importance of mesoscale eddies in 
> shaping particulate matter composition. Based on these results, we can infer the 
> biological communities and biogeochemical processes driven by mesoscale eddies at 
> both spatial and temporal scales. Viewing this complex dynamic process from the 
> chemical molecular scale is a perspective that sets this article apart from other 
> traditional methods in examining the changes in this system.
This is an accurate assessment of the manuscript.

> However, I still have some questions.
> Currently, the process of the entire system seems to be predictable. Although the 
> author has introduced and discussed some gaps, especially from the perspective of 
> chemical composition, and this article provides detailed data and reliable analysis, 
> it is not prominent enough from the perspective of the entire system. Perhaps the 
> transformation from space to time and the use of chemical composition to observe the 
> aging process of vortices is a great highlight. It is very difficult to focus on 
> scale eddies and the ecosystem evolution driven by them using traditional methods. 
> If that's the case, the structure and writing style of the paper may need to be 
> adjusted.
> This article focuses on the differences in chemical composition to understand this 
> process, but I always want to connect it with specific biological processes, such 
> as structural changes in biological communities, primary production rates, 
> zooplankton grazing rates, or bacterial production and respiration rates etc. 
> The author also mentioned these. I don't know if the author can add more relevant 
> evidence or discussion to allow readers to connect with these specific biological 
> groups and/or processes. I know it's difficult to integrate these data together, 
> but relevant discussions are possible.
> We already know that mesoscale eddies are complex and dynamic multidimensional 
> processes. The three-dimensional space combined with the temporal evolution of 
> ecosystem structure and function. The changes in the density surface driven by 
> physics during this process lead to the coupling reaction of physics chemistry 
> biology, which is the main thread. In fact, many other physical biological coupling 
> processes are similar, such as upwelling, internal waves/tides. Using the author's 
> method, I think similar situations can also be seen in such processes.
We have expanded the Discussion to touch on the concerns raised by this reviewer,
adding section 5.3 (starting at line 624) to respond to this directly.
Specifically, we have mentioned possible links between the increased sulfonium
compounds in the eddy and DMS production (lines 605-609), a hypothesis about the reason why
increased cyclonic biomass was not matched by increased export (lines 642-651), and extensive
references to additional manuscripts with relevant data throughout the section.

> Specific suggestions,
> Table 1, the format is not appropriate
Table 1 has been removed entirely.

> Figures 4, and 6 can be placed together.
While we agree that the data presented in Figures 4 and 6 can be placed side-by-side
for ease of comparison, we believe that keeping these figures in their separate
locations within the manuscript allows for better narrative flow (since the HEE
results are not discussed until several sections later) and graphical design (as
proper visualization of NMDS requires that the x- and y- axes share units, making 
it awkward to place them side-by-side).

> Reviewer 2 | 04 Oct 2024 | 12:20
> #1
> Kumler et al. report in their manuscript “Metabolites reflect variability introduced 
> by mesoscale eddies in the North Pacific Subtropical Gyre” on a very impressive 
> dataset, investigating the polar part of the meta-metabolome of a > 0.2 µm fraction 
> of the plankton. They investigate samples taken at the Meso-scope and HEE experiment 
> and discuss their data in light of the eddies found during sampling.
> It is clearly a high-quality study that opens new perspectives. Publication after 
> revision is highly recommended.
This is an accurate assessment of the manuscript.

> Sampling:
> The metabolome at different depths can be strongly affected by the sampling time. 
> Day / night variations are substantial, mainly due to diel vertical migration but 
> also to photosynthetic state of the phytoplankton. Discussing the results sampling 
> times during the transect and during sampling of the HEE should be specified and 
> statistical analysis on the impact (or lack of impact) on the data should be provided.
We have added the time of day as an explanatory factor to our analysis and reported the
new values from the PERMANOVA in the text (Lines 221-224, 263-267 and 522-525) and figures. 
We do see diel differences at the surface and DCM during the eddy transect and these have 
been reported, with the 175 meter differences more likely to be due to chance and very little
diel difference in the HEE samples at all. We have additionally added a supplemental
figure identifying a few select compounds for which diel effects are clearly visible
(trehalose and sucrose) alongside other highly abundant compounds which show no diel
effects.

> Analytics:
> The high-quality analytics is appreciated. It should be verified if the variations 
> between biological replicates could be caused by the order of measurements in the 
> sequence. If >100 samples of 15 min are run in a row there could be a drift that would 
> be hidden because of the randomization. It would be easy to have a look at the outliers 
> (e.g. Fig. 2 lower panels) are measured early / late during the sequence.
We have double-checked that our normalization to internal standards removed any
analytical drift and can confirm that while significant drift occurred over
the course of the eddy transect sample run, the normalization removed this
effect and is not a major factor in the outliers visible in Fig. 2. (Lines 193-195)

> Blanks are mentioned but it is not shown how these were used for the data analysis.
We have added a note about how we used the methodological blanks to visually assess 
whether sample peak areas were above the background signal. (Lines 187-189)

> In line 142 the method for the evaluation of the matrix effect is mentioned, but no 
> results are given.
The Boysen et al. (2018) paper details the method for identifying the magnitude of the 
matrix effect by using a full-strength and water-diluted pooled sample and we believe 
that the reader is best suited to the detailed analysis presented there than a
brief review of it in this manuscript. We have confirmed that references to
this method are cited immediately after its mention here.

> The experimental could be clearer in line 132 ff “The samples were then bead-beaten 
> three times, followed by triplicate extraction of the aqueous layer...” Was the 
> organic layer collected or was the aqueous layer collected (as suggested by the rest 
> of the sentence?). Was phase separation occurring spontaneously or was a 
> centrifugation required? Was separation done by pipetting?
We have added additional text to this line to clarify the requested points of 
confusion. (Lines 135-140)

> Line 177: I would rather state “targeted analytical methods”. Metabolomics is by 
> definition rather untargeted.
We have restructured the headings for this section of the methods and added
additional detail to respond to this concern.

> Line 185: The data are discussed quantitatively but for some compounds no standard 
> was available. The claim that results were normalized “to the best match standard”. 
> How was this decided? Are the different response-factors in the MS considered?
As above, the Boysen et al. (2018) paper describes how internal standards are matched
to compounds for which no authentic standard was available based on the difference
in their response factor between a full-strength and half-strength pooled sample.
The only data presented quantitatively (i.e. as concentrations instead of relative
or absolute peak area) are those for which an authentic standard was available,
ensuring that each quantified compound's response factor was known. For untargeted
molecular features for which no authentic standard was available, data is only
discussed explicitly in peak area space or as a relative deviation between stations/
depths. We have also added links to the Github where the R code used to perform
these matching and normalization steps is available. (Lines 195-200 and 206-209)

> Chapter 4.1.1 It is strange to read an entire chapter on an outlier. It is really 
> one measurement that could be easily contaminated (e.g. one single zooplankton 
> organism on the filter would change everything) and I would not recommend to write 
> an entire chapter about the metabolome, risking to discuss artifacts and not data 
> important for the large picture. Is table 1 required.
We appreciate the reviewer's identification of this as an unhelpful distraction
from the main narrative and have removed chapter 4.1.1 entirely.

> Minor concerns:
> Could the CTD data be given in the supporting material, this would provide a view of 
> the eddies in a depth profiles.
The CTD data has been added to the Supplemental Tables in 3 forms - first, containing
a list of all the stations and their locations with the number of bottles fired. Second,
with all the associated metadata compiled from Barone 2022. Finally, the values used
in the manuscript's analysis after interpolation to the locations in space and time
from which metabolite samples were collected.

> Software: Give the version number of the software used (e.g. Proteowizard, msConvert, 
> XCMS...)
Done. Software versions for Proteowizard, Skyline, MSDIAL, R, XCMS, and vegan have been
added to the manuscript. (Sections 3.6 and 3.7)

> Statistics: It would facilitate better interpretation if the statistical method used 
> for comparisons would be mentioned in the respective sections of results.
We have tried to add statistical methods used for comparisons in each section of the
results where applicable in this revision, largely annotating the places in which 
R-squared and p-values were obtained from PERMANOVAs. We have also added the sample
size to the Statistics section of the Results. (Lines 224-227)

> Line 239 Typo r = 0.697
We have switched to using a 0 in front of Pearson's correlation coefficient throughout
the manuscript.

> Line 266: From the figures in the supporting material, it does not become evident that 
> Particulate caron and metabolite concentrations are connected. There could be nested 
> effects due to the selection of data points.
It is unclear what the reviewer means by the selection of data points. We are aware that
the strong trend between particulate carbon and metabolite concentration is largely due
to the effect of depth and that was our inspiration for performing both the overall
regression and independent regressions at each depth.